### 2015-04-08 concl.tex

This paper presents \name, a shallow Haskell extension for declarative
programming over eventually consistent data stores. The key idea underlying
\name's design is the automatic classification of fine-grained consistency
contracts on operations and distributed transactions with respect to the
consistency and isolation levels offered by the store. Our contract language
is carefully crafted from a decidable subset of first-order logic, enabling
the use of automated verification tools to discharge the proof obligations
associated with contract classification. We realize an instantiation of
\name on top of off-the-shelf distributed store, Cassandra, and illustrate
the benefit of fine-grained contract classification by implementing and
evaluating several scalable applications.

>>>>

This paper presents \name, a shallow Haskell extension for declarative
programming over \ecds. The key idea underlying \name's design is the automatic
classification of fine-grained consistency contracts on operations and
distributed transactions with respect to the consistency and isolation levels
offered by the store. Our contract language is carefully crafted from a
decidable subset of first-order logic, enabling the use of automated
verification tools to discharge the proof obligations associated with contract
classification. We realize an instantiation of \name on top of off-the-shelf
distributed store, Cassandra, and illustrate the benefit of fine-grained
contract classification by implementing and evaluating several scalable
applications.

### 2015-04-08 contract-language.tex
Minor edit

\begin{figure}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{rclcl}
\multicolumn{5}{c}{
  {x,y,\cureff} \in \mathtt{EffVar} \qquad \qquad
  {\sf Op} \in \mathtt{OperName}
}\\
\cv 		& \in & \mathtt{Contract} 	& \coloneqq & \forall (x : \tau).\cv
        \ALT \forall x.\cv \ALT \pi \\
\tau		& \in	& \mathtt{EffType}	& \coloneqq &  {\sf Op}
        \ALT \tau \vee \tau \\
\pi			&	\in & \mathtt{Prop} & \coloneqq & \true \ALT R(x,y)
        \ALT \pi \vee \pi \\
			  & 		&	 &  \ALT & \pi \wedge \pi \ALT \pi \Rightarrow \pi \\
R				& \in & \mathtt{Relation}	& \coloneqq & \visZ \ALT \soZ
        \ALT \sameobjZ \ALT R^+ \\
				&			&	 &  \ALT & R \cup R \ALT R \cap R \\
\end{array}
\end{smathpar}
\caption{Contract language.}
\label{fig:contract-lang}
\end{figure}

@@

\begin{figure}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

\begin{itemize}

@@

  \vspace{-1mm}
  \begin{cmathpar}
  \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{cmathpar}

@@

  \vspace{-1mm}
  \begin{cmathpar}
  \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{cmathpar}

\item \textbf{Strong consistency}: Strongly consistent operations may block
  indefinitely under network partitions. An example is the total-order
  contract on \cf{withdraw} operation. The corresponding store semantics is
	captured by the $\scc$ contract definition:

  \vspace{-1mm}
  \begin{cmathpar}
  \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
  \end{cmathpar}
\end{itemize}

@@

\newcommand{\DDe}[1]{#1}
\begin{figure}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{cc}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  &

@@

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{itemize}
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

\subsection{Generality of Contracts}

It is important to note that our contract language provides a generic way to
capture application-level consistency properties and is not tied to a
particular store semantics. In particular, the same application-level contracts
can easily be mapped to a different store with a varied consistency lattice. To
illustrate this, let us consider the consistency lattice proposed by Terry et
al.~\cite{Session} based on session guarantees. Terry et al. propose the
following four incomparable session guarantees, whose semantics is captured in
the contracts below:

\vspace{-2mm}
\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{rcl}
\R{Read Your Writes (RYW)} & \coloneqq & \forall a,b. ~\soo{a}{b} \Rightarrow \vis{a}{b} \\
\R{Monotonic Reads (MR)} & \coloneqq & \forall a,b,c. ~\vis{a}{b} \wedge \soo{b}{c} \\
 & & \quad \Rightarrow \vis{a}{c} \\
\R{Monotonic Writes (MW)} & \coloneqq & \forall a,b,c. ~\soo{a}{b} \wedge \vis{b}{c} \\
 & & \quad \Rightarrow \vis{a}{c} \\
\R{Writes Follow Reads (WFR)} & \coloneqq & \forall a,b,c,d. ~\vis{a}{b} \wedge \vis{c}{d} \\
 & & \wedge (\sooZ ~\cup =)(b,c) \Rightarrow \vis{a}{d}
\end{array}
\end{smathpar}

Here, the consistency level for an operation in this scheme is any combination
of the above guarantees. It is easy to see that there are 24 (4!) combinations,
which form a partially ordered consistency lattice with a height of 4.
Classifying a contract under this scheme amounts to determining the weakest
combination under which the contract will hold, and entails 24 comparisons at
worst. Under this scheme, \cf{deposit} operations does not need any guarantees,
\cf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge \R{WFR}
\Rightarrow \cv_{gb}$), and \cf{withdraw} cannot be satisfied ($\Delta \vdash
\R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow \cv_{w}$).

@@

\begin{smathpar}
\renewcommand{\arraystretch}{1.2}
\begin{array}{lclcl}
{\it op} 	& \in & \mathtt{Operation} \\
{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
\end{array}
\end{smathpar}

@@

\begin{smathpar}
  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E',\sigma \pll \Sigma}
\end{smathpar}

@@

The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
$\ecc$ contracts in all well-formed executions, then the same store, extended
with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
all well-formed \name contracts. The proof of the theorem is given in the
supplementary material\footnote{The supplementary material also provides the
concrete reduction rules for enforcing the consistency classes.}.

>>>>

\begin{figure}
\vspace{-7pt}
\begin{smathpar}
\begin{array}{rclcl}
\multicolumn{5}{c}{
  {x,y,\cureff} \in \mathtt{EffVar} \qquad \qquad
  {\sf Op} \in \mathtt{OperName}
}\\
\cv 		& \in & \mathtt{Contract} 	& \coloneqq & \forall (x : \tau).\cv
        \ALT \forall x.\cv \ALT \pi \\
\tau		& \in	& \mathtt{EffType}	& \coloneqq &  {\sf Op}
        \ALT \tau \vee \tau \\
\pi			&	\in & \mathtt{Prop} & \coloneqq & \true \ALT R(x,y)
        \ALT \pi \vee \pi \\
			  & 		&	 &  \ALT & \pi \wedge \pi \ALT \pi \Rightarrow \pi \\
R				& \in & \mathtt{Relation}	& \coloneqq & \visZ \ALT \soZ
        \ALT \sameobjZ \ALT R^+ \\
				&			&	 &  \ALT & R \cup R \ALT R \cap R \\
\end{array}
\end{smathpar}
\caption{Contract language.}
\label{fig:contract-lang}
\end{figure}

@@

\begin{figure}
\vspace{-7pt}
\begin{smathpar}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

\begin{itemize}[noitemsep]

@@

  \vspace{-1.5em}
  \begin{smathpar}
  \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

@@

  \vspace{-1.5em}
  \begin{smathpar}
  \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
  \end{smathpar}

\item \textbf{Strong Consistency}: Strongly consistent operations may block
  indefinitely under network partitions. An example is the total-order
  contract on \cf{withdraw} operation. The corresponding store semantics is
	captured by the $\scc$ contract definition:

  \vspace{-1.5em}
  \begin{smathpar}
  \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
  \end{smathpar}
\end{itemize}
\vspace{-1em}

@@

\newcommand{\DDe}[1]{#1}
\begin{figure}
\vspace{-10pt}
\begin{smathpar}
\begin{array}{cc}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  &

@@

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\vspace{-3mm}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{itemize}[noitemsep]
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

@@

@@

\vspace{-1em}
\begin{smathpar}
\begin{array}{lclcl}
{\it op} 	& \in & \mathtt{Operation} \\
{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
\end{array}
\end{smathpar}

@@

\vspace{-1.7em}
\begin{smathpar}
  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E',\sigma \pll \Sigma}
\end{smathpar}

@@

\noindent The theorem states that if a data store correctly enforces $\scc$,
$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
store, extended with the classification scheme shown in
Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
proof of the theorem is given below:

\begin{proof}
  Hypothesis:
  \begin{smathpar}
  \begin{array}{cr}
    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E', \sigma \pll \Sigma} & H\npp\\
    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
  \end{array}
  \end{smathpar}
  Since $\tau$ is the contract class of $\cv$, by inversion, we have
  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
  \end{array}
  \end{smathpar}
   Since $\eff$ denotes new effect, it is a fresh variable that does
   not occur free in $\Delta$. From $H3$, after instantiating bound
   $\cureff$ with $\eff$, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
      & H\npp\\
  \end{array}
  \end{smathpar}
  Due to the soundness of natural deduction for first-order logic,
  $H4$ implies that for all models $\mathcal{M}$ such that
  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \E' \models \Delta & H\npp\\
  \end{array}
  \end{smathpar}
  Proof follows from $H1$, $H4$, and $H3$.
  \hfill \qed
\end{proof}

It is important to note that Theorem~\ref{thm:classification-sound}
does not ascribe any semantics to the reduction relation
($\xrightarrow{}$). As such, it makes no assumptions about how the
store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
guarantees. The specific implementation strategy is determined by the
operational semantics of the store, which \emph{defines} the reduction
relation for that particular store. The following section describes
operational semantics of the store used by the \name implementation.

### 2015-04-08 related.tex

Since eventual consistency alone is insufficient to build correct applications,
several systems~\cite{Bayou,Pileus,RedBlue} propose a lattice of stronger
consistency levels. Similarly, traditional database processing
systems~\cite{Berenson95} and their replicated variants~\cite{BailisHAT}
propose weaker isolation levels for performance. In these systems, the onus is
on the developer to choose the correct consistency(isolation) level for
operations(transactions). \name relieves the developer of this burden, and
instead expects contracts expressing declarative visibility requirements.

>>>>

Since eventual consistency alone is insufficient to build correct applications,
several systems~\cite{Bayou,Pileus,RedBlue} propose a lattice of stronger
consistency levels. Similarly, traditional database processing
systems~\cite{Berenson95} and their replicated variants~\cite{BailisHAT}
propose weaker isolation levels for performance. In these systems, the onus is
on the developer to choose the correct consistency(isolation) level for
operations(transactions). \name relieves the developer of this burden, and
instead expects programmers to declare application-specific consistency
requirements as first-order formulas. In this regard, \name is similar to Indigo
\cite{indigo}, which requires programmers to declare application-specific
invariants that are expected to be met, along with pre- and post-conditions on
operations as formulas in first-order logic. Indigo then performs a static
analysis on formulas to identify $I$-offender sets - sets of operations,
which, when performed concurrently, result in violation of one or more of the
stated invariants. For each offending set of operation, if programmer chooses 
invariant-violation avoidance over violation repair, the system employs various
techniques, such as escrow reservation, to ensure that the offending set of
operations are never run concurrently.

While serializing operations is sufficient to ensure that the application state
never violates invariants (e.g: \cf{balance $\ge$ 0}), there exist a range of
desirable application-level properties, which cannot be captured in Indigo's
specification language, or for which serialization is an overkill. For instance,
it is neither efficient nor straightforward in Indigo to declare and enforce a
condition to prevent the missing update anomaly
(Figure~\ref{fig:missingUpdateAnomaly}) in the bank account application. The
only way is to encode specifications of all bank account operations such that
every pair of operations are identified as $I$-offending, for some invariant
($I$), resulting in them being serialized. In contrast, \name's contract
language allows for precise encoding of the desired property, and the contract
classification scheme ensures that the it can be enforced without the need to
serialize operations. Furthermore, Indigo also does not distinguish between
various transaction isolation guarantees, which \name does.
